---
title: "TSA Forecasting_report_top5"
output: html_document
date: "2026-04-23"
---

> Summary of results: The TBATS model achieved the best performance due to its ability to capture multiple seasonal patterns in daily electricity demand. Other models incorporating exogenous variables did not significantly improve accuracy.
Model performace rank from best to worst: 
1. TBATS
2. SARIMAX
3. STLM
4. MSTL_ETS
5. ARIMA + Fourier

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(readxl)
library(forecast)
```

## 1. Load Dataset

#### Preprocessing

```{r}
load <- read_excel("load.xlsx")
head(load)
```

#### Convert Hourly to Daily

```{r}
daily_load <- load %>%
  rowwise() %>%
  mutate(daily_load = mean(c_across(starts_with("h")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(date, daily_load)

daily_load$date <- as.Date(daily_load$date)

```

#### Train-Validation Split

```{r}
train_data <- daily_load %>%
  filter(date <= as.Date("2010-06-30"))

valid_data <- daily_load %>%
  filter(date >= as.Date("2010-07-01") &
         date <= as.Date("2010-07-31"))

train_msts <- msts(
  train_data$daily_load,
  seasonal.periods = c(7, 365),
  start = c(2005, 1)
)

train_ts <- ts(train_data$daily_load,
               start = c(2005,1),
               frequency = 365)

valid_msts <- msts(
  valid_data$daily_load,
  seasonal.periods = c(7, 365),
  start = c(2010, 182)
)

valid_ts <- ts(valid_data$daily_load,
               start = c(2010,7),
               frequency = 365)
```

#### Decomposition

```{r}
decomp <- stl(train_ts, s.window = "periodic")
autoplot(decomp) # seasonal
```

## 2. Model 1 - TBATS (Baseline Model)

```{r}
fit_tbats <- tbats(train_msts)
forecast_tbats <- forecast(fit_tbats, h = length(valid_msts))

autoplot(forecast_tbats) +
  autolayer(valid_msts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```

```{r}
summary(fit_tbats)
accuracy(forecast_tbats, valid_msts)
```

## 3. Temperature Dataset

#### Preprocessing

```{r}
# load data
temperature <- read_excel("temperature.xlsx")
head(temperature)
```

```{r}
# convert hourly to daily
temp_daily <- temperature %>%
  mutate(date = as.Date(date, format="%d%b%Y")) %>%
  group_by(date) %>%
  summarise(across(-hr, mean, na.rm = TRUE))

# split train data
temp_train <- temp_daily %>%
  filter(date <= as.Date("2010-06-30"))

# convert to time series
temp_train_ts <- ts(
  temp_train[,-1],
  start = c(2005,1),
  frequency = 365
)
```

#### Determining each station's level of correlations with load

```{r}
cors <- cor(train_ts, temp_train_ts, use="complete.obs")
sort(cors, decreasing = TRUE)
```

-   **All correlations are around -0.41 to -0.47. That means:**

    -   **Temperature and load are moderately negatively correlated.**

    -   **The stations behave very similarly (only \~0.05 difference between best and worst).**

```{r}
# average all station data since their correlations are similar
temp_avg_ts <- ts(
  rowMeans(temp_train_ts, na.rm = TRUE),
  start = start(temp_train_ts),
  frequency = frequency(temp_train_ts)
)
```

```{r}
ccf(train_ts, temp_avg_ts, lag.max = 7)
```

**Investigating correlations between average temperature and a lagged load series showed no** **improvement in correlations.**

## 4. Humidity Dataset

#### Preprocessing

```{r}
# load data
humidity <- read_excel("relative_humidity.xlsx")
head(humidity)
```

```{r}
# convert hourly to daily
humi_daily <- humidity %>%
  mutate(date = as.Date(date, format="%d%b%Y")) %>%
  group_by(date) %>%
  summarise(across(-hr, mean, na.rm = TRUE))

# split train data
humi_train <- humi_daily %>%
  filter(date <= as.Date("2010-06-30"))

# convert to time series
humi_train_ts <- ts(
  humi_train[,-1],
  start = c(2005,1),
  frequency = 365
)
```

#### Determining each station's level of correlations with load

```{r}
cors_humi <- cor(train_ts, humi_train_ts, use="complete.obs")
sort(cors_humi, decreasing = TRUE)
```

**Relative humidity have an even weaker negative correlations with load than the temperature measurements.**

**I'll average the top 5 stations with a correlation of 0.225 and larger in absolute value.**

```{r}
top5 <- order(abs(cors), decreasing = TRUE)[1:5]

humi_avg_ts <- ts(
  rowMeans(humi_train_ts[, top5], na.rm = TRUE),
  start = start(humi_train_ts),
  frequency = frequency(humi_train_ts)
)
```

```{r}
ccf(train_ts, humi_avg_ts, lag.max = 7)
```

**Investigating correlations between average temperature and a lagged load series showed no** **improvement in correlations.**

## 5. Model 2 - SARIMAX Model with Temperature and Humidity

```{r}
ts_train <- ts(
  train_data$daily_load,
  start = c(2005, 1),
  frequency = 365
)

valid_ts <- ts(
  valid_data$daily_load,
  start = c(2010, 182),
  frequency = 365
)

xreg_train <- cbind(
  temp = train_data$temp,
  humi = train_data$humidity
)

xreg_valid <- cbind(
  temp = valid_data$temp,
  humi = valid_data$humidity
)

fit_sarimax <- auto.arima(
  ts_train,
  xreg = xreg_train,
  seasonal = TRUE
)
```

```{r}
fc_sarimax <- forecast(
  fit_sarimax,
  xreg = xreg_valid,
  h = nrow(valid_data)
)

autoplot(fc_sarimax) +
  autolayer(valid_ts, series = "Actual") +
  ylab("electricity load")
```
```{r}
summary(fit_sarimax)
accuracy(fc_sarimax, valid_ts)
```

## 6. Model 3 - ARIMA + Fourier Terms

```{r}
best_aicc <- Inf
best_fit <- NULL
best_K <- NULL

for(k1 in 1:3){        # weekly (max 3)
  for(k2 in 1:10){     # yearly (safe range)
    
    xreg <- fourier(train_msts, K = c(k1, k2))
    
    fit <- auto.arima(
      train_msts,
      xreg = xreg,
      seasonal = FALSE
    )
    
    if(fit$aicc < best_aicc){
      best_aicc <- fit$aicc
      best_fit <- fit
      best_K <- c(k1, k2)
    }
  }
}

best_K
```

```{r}
K <- c(3,1)
# 3 Fourier pairs for weekly, 5 for yearly

fourier_train <- fourier(train_msts, K = K)

fit_arima_fourier <- auto.arima(
  train_msts,
  xreg = fourier_train,
  seasonal = FALSE
)

fourier_future <- fourier(train_msts, K = K, h = length(valid_msts))

forecast_arima_fourier <- forecast(
  fit_arima_fourier,
  xreg = fourier_future,
  h = length(valid_msts)
)

autoplot(forecast_arima_fourier) +
  autolayer(valid_msts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```

```{r}
summary(fit_arima_fourier)
accuracy(forecast_arima_fourier, valid_msts)
```

## 8. Model 4 - MSTL + ETS

```{r}
fit_mstl <- mstl(train_msts)

fit_ets <- ets(seasadj(fit_mstl))

fc_mstl <- forecast(fit_ets, h = length(valid_msts))

autoplot(fc_mstl) +
  autolayer(valid_msts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```

```{r}
summary(fit_ets)
accuracy(fc_mstl, valid_msts)
```

## 9. Model 5 - STLM

```{r}
fit_stlm <- stlm(train_msts, method = "ets")

fc_stlm <- forecast(fit_stlm, h = length(valid_msts))

autoplot(fc_stlm) +
  autolayer(valid_msts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```
```{r}
summary(fit_stlm)
accuracy(fc_stlm, valid_msts)
```
> Model performaces interpretation: 
Among the five models evaluated, The baseline TBATS model consistently achieved the lowest RMSE, MAE, and MAPE on the test set, indicating best predictive accuracy.
In contrast, The other 4 exhibited higher forecast errors, with the ARIMA + Fourier model performing the worst across all metrics.
Notably, the Theil’s U statistic for Model 1 is closest to 1, suggesting that its performance is competitive relative to a naïve benchmark, while other models significantly underperform.

## Export

#### Export Function

```{r}
export_submission <- function(forecast_obj, filename) {
  
  # extract forecast values
  fc_values <- as.numeric(forecast_obj$mean)
  
  # create July 2011 dates
  dates <- seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day")
  
  # keep only first 31 forecast values
  df <- data.frame(
    date = dates,
    load = fc_values[1:length(dates)]
  )
  
  # write csv
  write.csv(df, filename, row.names = FALSE)
}
```

#### TBATS baseline model

```{r}
export_submission(forecast_tbats, "tbats_submission.csv")
```

#### SARIMAX model

```{r}
export_submission(fc_arimax, "SARIMAX_submission.csv")
```

#### Fourier Terms

```{r}
export_submission(forecast_arima_fourier,"ARIMA_Fourier.csv")
```

#### MSTL

```{r}
export_submission(fc_mstl,"MSTL_ETS.csv")
```

#### STLM

```{r}
export_submission(fc_stlm,"STLM.csv")
```
