---
title: "TSA Forecasting_1"
output: html_document
date: "2026-03-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(readxl)
library(forecast)
```

## 1. Load Dataset

#### Preprocessing

```{r}
load <- read_excel("load.xlsx")
head(load)
```

#### Convert Hourly to Daily

```{r}
daily_load <- load %>%
  rowwise() %>%
  mutate(daily_load = mean(c_across(starts_with("h")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(date, daily_load)

daily_load$date <- as.Date(daily_load$date)

```

#### Train-Validation Split

```{r}
train_data <- daily_load %>%
  filter(date <= as.Date("2010-06-30"))

valid_data <- daily_load %>%
  filter(date >= as.Date("2010-07-01") &
         date <= as.Date("2010-07-31"))

train_msts <- msts(
  train_data$daily_load,
  seasonal.periods = c(7, 365),
  start = c(2005, 1)
)

train_ts <- ts(train_data$daily_load,
               start = c(2005,1),
               frequency = 365)

valid_msts <- msts(
  valid_data$daily_load,
  seasonal.periods = c(7, 365),
  start = c(2010, 182)
)

valid_ts <- ts(valid_data$daily_load,
               start = c(2010,7),
               frequency = 365)
```

## 2. TBATS (Baseline Model)

```{r}
fit_tbats <- tbats(train_msts)
forecast_tbats <- forecast(fit_tbats, h = length(valid_ts))

autoplot(forecast_tbats) +
  autolayer(valid_ts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```

## 3. Temperature Dataset

#### Preprocessing

```{r}
# load data
temperature <- read_excel("temperature.xlsx")
head(temperature)
```

```{r}
# convert hourly to daily
temp_daily <- temperature %>%
  mutate(date = as.Date(date, format="%d%b%Y")) %>%
  group_by(date) %>%
  summarise(across(-hr, mean, na.rm = TRUE))

# split train data
temp_train <- temp_daily %>%
  filter(date <= as.Date("2010-06-30"))

# convert to time series
temp_train_ts <- ts(
  temp_train[,-1],
  start = c(2005,1),
  frequency = 365
)
```

#### Determining each station's level of correlations with load

```{r}
cors <- cor(train_ts, temp_train_ts, use="complete.obs")
sort(cors, decreasing = TRUE)
```

-   **All correlations are around -0.41 to -0.47. That means:**

    -   **Temperature and load are moderately negatively correlated.**

    -   **The stations behave very similarly (only \~0.05 difference between best and worst).**

```{r}
# average all station data since their correlations are similar
temp_avg_ts <- ts(
  rowMeans(temp_train_ts, na.rm = TRUE),
  start = start(temp_train_ts),
  frequency = frequency(temp_train_ts)
)
```

```{r}
ccf(train_ts, temp_avg_ts, lag.max = 7)
```

**Investigating correlations between average temperature and a lagged load series showed no** **improvement in correlations.**

## 4. Humidity Dataset

#### Preprocessing

```{r}
# load data
humidity <- read_excel("relative_humidity.xlsx")
head(humidity)
```

```{r}
# convert hourly to daily
humi_daily <- humidity %>%
  mutate(date = as.Date(date, format="%d%b%Y")) %>%
  group_by(date) %>%
  summarise(across(-hr, mean, na.rm = TRUE))

# split train data
humi_train <- humi_daily %>%
  filter(date <= as.Date("2010-06-30"))

# convert to time series
humi_train_ts <- ts(
  humi_train[,-1],
  start = c(2005,1),
  frequency = 365
)
```

#### Determining each station's level of correlations with load

```{r}
cors_humi <- cor(train_ts, humi_train_ts, use="complete.obs")
sort(cors_humi, decreasing = TRUE)
```

**Relative humidity have an even weaker negative correlations with load than the temperature measurements.**

**I'll average the top 5 stations with a correlation of 0.225 and larger in absolute value.**

```{r}
top5 <- order(abs(cors), decreasing = TRUE)[1:5]

humi_avg_ts <- ts(
  rowMeans(humi_train_ts[, top5], na.rm = TRUE),
  start = start(humi_train_ts),
  frequency = frequency(humi_train_ts)
)
```

```{r}
ccf(train_ts, humi_avg_ts, lag.max = 7)
```

**Investigating correlations between average temperature and a lagged load series showed no** **improvement in correlations.**

## 5. SARIMAX Model

```{r}
xreg <- cbind(temp_avg_ts, humi_avg_ts)

fit_arimax <- auto.arima(
  train_ts,
  xreg = xreg,
  seasonal = TRUE
)
```

```{r}
fc_arimax <- forecast(fit_arimax, xreg = xreg, h = length(valid_ts))

autoplot(fc_arimax) +
  autolayer(valid_ts, series = "Actual", PI = FALSE) +
  xlim(2010, 2010.6) +
  ylab("electricity load")
```

## 10. Export

#### Export Function

```{r}
export_submission <- function(forecast_obj, filename) {
  
  # extract forecast values
  fc_values <- as.numeric(forecast_obj$mean)
  
  # create July 2011 dates
  dates <- seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day")
  
  # keep only first 31 forecast values
  df <- data.frame(
    date = dates,
    load = fc_values[1:length(dates)]
  )
  
  # write csv
  write.csv(df, filename, row.names = FALSE)
}
```

#### TBATS baseline model

```{r}
export_submission(forecast_tbats, "tbats_submission.csv")
```

#### SARIMAX model

```{r}
export_submission(fc_arimax, "SARIMAX_submission.csv")
```
